|

### 2021-07-31 prova_knit_finale.Rmd
sto facendo implementations SV model

Finally we 
\begin{algorithm} LWF for Random Walk plus Noise Model
\begin{itemize}
\item Initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$, $(\alpha^{(1)},...,\alpha^{(N)})$ from $N(e_{\alpha},v_\alpha)$ and
$(\beta^{(1)},...,\beta^{(N)})$ from $N(e_{\beta},v_\beta)$ $({\tau^{2}}^{(1)},...,{\tau^{2}}^{(N)})$ from $IG(\nu_{0}/2,\nu_{0}\lambda/2)$. Set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$ and chose the appropriate value for $\delta$. Therefore $\psi^{(i)}=({\alpha}^{(i)},{\beta}^{(i)},{\tau^{2}}^{(i)})$, and
$\hat{\pi}_{0}=p(x_{0}|y_{0})=\sum_{i=1}^{N}w_{0}^{(i)}\delta_{(x_{0}^{(i)},\psi^{(i)})}$
\item For $t=1,...,n$:.
\begin{enumerate}
\item Compute $\overline{\psi}=E_{\hat{\pi}_{t-1}}(\psi)$ and $\Sigma=V_{\hat{\pi}_{t-1}}(\psi)$. For $i=1,...,N$, set
\begin{align*}
m^{(i)} & = a\psi^{(i)}+(1-a)\hat{\psi} \\
\hat{x}_{t}^{(i)} & = E(x_{t}|x_{t-1}=x_{t-1}^{(i)},\psi=\psi^{(k)})
\end{align*}
\item For $k=1,...,N$:
\begin{itemize}
\item Draw $I_{k}$, with $P(I_{k}=i) \propto w_{t-1}^{(i)}f_{N}(y_{t}|0,e^{\frac{\hat{x}_{t}^{(i)}}{2}}) $
\item Draw $\psi^{(k)}$ from $N(m^{(I_{k})},h^{2}\Sigma)$
\item Draw $x_{t}^{(k)}$ from $N(\alpha^{(k)}+\beta^{(k)}x_{t-1}^{(I_{k})},\tau^{2})$
\item Set  $\tilde{w}_{t}^{(k)} = \frac{f_{N}(y_{t}|0,e^{\frac{x_{t}^{(k)}}{2}})}{f_{N}(y_{t}|0,e^{\frac{\hat{x}_{t}^{(I_{k})}}{2}})}$
\end{itemize}
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $\hat{\pi}_{t}=p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{(x_{t}^{(i)},\psi^{(i)})}$
\end{enumerate}
\end{itemize}
\end{algorithm}

>>>>

Finally we present a Liu and West Filter for the case of parameters $\alpha$, $\beta$ and $\tau^{2}$ unknown.

\begin{algorithm} LWF for Stochastic Volatility Model
\begin{itemize}
\item Initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$, $(\alpha^{(1)},...,\alpha^{(N)})$ from $N(e_{\alpha},v_\alpha)$ and
$(\beta^{(1)},...,\beta^{(N)})$ from $N(e_{\beta},v_\beta)$ $({\tau^{2}}^{(1)},...,{\tau^{2}}^{(N)})$ from $IG(\nu_{0}/2,\nu_{0}\lambda/2)$. Set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$ and chose the appropriate value for $\delta$. Therefore $\psi^{(i)}=({\alpha}^{(i)},{\beta}^{(i)},{\tau^{2}}^{(i)})$, and
$\hat{\pi}_{0}=p(x_{0}|y_{0})=\sum_{i=1}^{N}w_{0}^{(i)}\delta_{(x_{0}^{(i)},\psi^{(i)})}$
\item For $t=1,...,n$:.
\begin{enumerate}
\item Compute $\overline{\psi}=E_{\hat{\pi}_{t-1}}(\psi)$ and $\Sigma=V_{\hat{\pi}_{t-1}}(\psi)$. For $i=1,...,N$, set
\begin{align*}
m^{(i)} & = a\psi^{(i)}+(1-a)\overline{\psi} \\
\hat{x}_{t}^{(i)} & = E(x_{t}|x_{t-1}=x_{t-1}^{(i)},\psi=\psi^{(k)})
\end{align*}
\item For $k=1,...,N$:
\begin{itemize}
\item Draw $I_{k}$, with $P(I_{k}=i) \propto w_{t-1}^{(i)}f_{N}(y_{t}|0,e^{\frac{\hat{x}_{t}^{(i)}}{2}}) $
\item Draw $\psi^{(k)}$ from $N(m^{(I_{k})},h^{2}\Sigma)$
\item Draw $x_{t}^{(k)}$ from $N(\alpha^{(k)}+\beta^{(k)}x_{t-1}^{(I_{k})},\tau^{2})$
\item Set  $\tilde{w}_{t}^{(k)} = \frac{f_{N}(y_{t}|0,e^{\frac{x_{t}^{(k)}}{2}})}{f_{N}(y_{t}|0,e^{\frac{\hat{x}_{t}^{(I_{k})}}{2}})}$
\end{itemize}
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $\hat{\pi}_{t}=p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{(x_{t}^{(i)},\psi^{(i)})}$
\end{enumerate}
\end{itemize}
\end{algorithm}

We are now ready to compare the performances of these four algorithms. We simulate a Stochastic Volatility process setting arbitrarily the following parameters: $\alpha = -0.1$, $\beta = 0.9$ and $\tau^{2}=0.01$. The resulting observation and state process is shown in the figure below.

```{r include=FALSE}
# Simulating the data
# -------------------
set.seed(12345)
n     =  50
alpha = -0.1
beta  =  0.9
tau2  =  0.01
tau   = sqrt(tau2)
y     = rep(0,n)
x     = rep(0,n)
x[1]  = alpha/(1-beta)
y[1]  = rnorm(1,0,exp(x[1]))
for (t in 2:n){
  x[t] = alpha + beta*x[t-1] + rnorm(1,0,tau)
  y[t] = rnorm(1,0,exp(x[t]))
}
alpha.true = alpha
beta.true  = beta
tau2.true  = tau2

timeframe<-c(1:length(y))
dfsv<-data.frame(timeframe,y,x)
```

```{r echo=FALSE, fig.align='center', fig.cap="Simulated Stochastic Volatility Model", fig.height=4, fig.pos='ht', fig.width=8, message=FALSE, warning=FALSE, results=F}
ggplot(dfsv,aes(x=timeframe))+
  geom_line(aes(y=y, linetype="Observations"))+
  geom_line(aes(y=x, linetype="True States"))+
  scale_linetype_manual(name="",
                      values=c("Observations" = 1,
                               "True States" = 3))+
  labs(x="Time",
       y="")+
  theme_bw()+
  theme(legend.direction = "horizontal", legend.position = "bottom", legend.key = element_blank(), 
        legend.background = element_rect(fill = "white", colour = "gray30")) +
  theme(plot.title = element_text(hjust = 0.5))
```
